Chemical potential of liquids and mixtures via Adaptive Resolution Simulation

### 2014-05-19 writeDown/pub.jcp/resubmission.tex
resubmission letter imporoved

\documentclass[12pt]{article}
\usepackage{german}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{pslatex}
\usepackage[]{color}
\usepackage{amsmath,amssymb,amsfonts,latexsym,graphicx}

\pagestyle{empty}
\parindent0cm
% \addtolength{\textheight}{2cm}
% \addtolength{\textwidth}{1cm}
\renewcommand{\sfdefault}{cmss}
\renewcommand{\familydefault}{\sfdefault}
\newcommand{\redc}[1]{{\color{red} #1}}            
\newcommand{\recheck}[1]{{\color{red} #1}}

\newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\dof}{{\textrm{DOF}}}
\newcommand{\AT}{{\textrm{{AT}}}}
\newcommand{\EX}{{\textrm{EX}}}
\newcommand{\CG}{{\textrm{CG}}}
\newcommand{\HY}{{\Delta}}
\newcommand{\rdf}{{\textrm{rdf}}}
\newcommand{\thf}{{\textrm{th}}}
\newcommand{\res}{{\textrm{rep}}}
\newcommand{\ext}{{\textrm{extra}}}
\newcommand{\exc}{{\textrm{exc}}}
\newcommand{\thermo}{{\textrm{Q}}}
\newcommand{\hadress}{{\textrm{H}}}
\newcommand{\dadress}{{\textrm{D}}}

\newcommand{\kopf}{\noindent
\raisebox{2.5cm}[0cm][0cm]
{
\parbox[c]{0.88\textwidth}
{
\begin{center}
\textsc{\large Institute for Mathematics, Freie Universit\"{a}t Berlin}\\
\textsc{ PD Dr. Luigi Delle Site}
\end{center}
}
}}
%\raisebox{2cm}[0cm][0cm]{\rule{\textwidth}{0.2mm}}}
\begin{document}

\kopf
\vspace{2cm}
\hfill\raisebox{1.5cm}[0cm][0cm]
{\parbox[t]{0.34\textwidth}
{e-mail:\\
{\small luigi.dellesite@fu-berlin.de}\\
Arnimallee 6\\
D-14195 Berlin\\
Tel:\ +49 (0)30 838 75775 \\
Fax:\ + 49 (0)30 838 75412 \\
}}

\vspace{3cm}
\noindent\raisebox{1.5cm}[0cm][0cm]
{\fbox{\parbox[t]{0.6\textwidth}
{Editor of {\it The Journal of Chemical Physics}\\
}}}
\vspace{1cm}

@@

\vspace{1cm}
\parskip 2ex
Dear Editor\\
We would like to thank both referees for the criticisms and suggestions.
We agree on several points, however tend to strongly disagree on some crucial ones.\\
Below please find a reply to each of the concerns raised by the referees and the
corresponding modification of the manuscript (where required).\\
Moreover one section is dedicated to comments that are for you and are properly labeled.\\

{\bf For the Editor:}

Regarding the comment of {\bf Reviewer 1}:

{\it  It is not easy to judge the relevance of this
  work because it might contain enough simulations, but not new information. In fact, its main contribution is a table, with comparison with reported data.}

In the previous paper published about one year ago in Physical Review X,
we have developed what we have {\bf openly} proposed in that paper as a {\bf very rigorous approach} where all the formulas were derived analytically and checked mathematically.

The whole theory is based on two assumptions:

(1) Thermodynamic limit for the \redc{atomistic and the coarse-grained region. (i.e. the number of molecules in both of the two
  regions are infinitely large, and the coarse-grained region is infinitely larger than the atomistic region)}

@@

In a simulation of course these conditions cannot be met at the ideal
level, 
thus the idea of this paper was that of treating real systems
in order to explore the numerical {\bf accuracy} that the formulas can
provide. Thirteen systems were studied with both AdResS and with the
thermodynamic integration, this {\it per se} is a major numerical
effort, \recheck{I would delete ``even for large computer facilities'', see the reason below}, however, the systems were
carefully chosen as representative of different classes of liquids and
mixtures characterized by different physics. The {\bf new information}
consists of thirteen values of the chemical potential which are
reproduced by AdResS, thus showing that the method is trustable for
such kind of calculations and that the hypothesis (1) and (2) above
can be easily met in \recheck{systems of relatively small sizes (see Table~II of the manuscript). Actually
  all simulations reported were performed on a 8-core workstation.}
\recheck{Comment from HAN: the thermodynamic limit does not imply extremely large system,
  rather, comparatively small systems that can be simulated on desktop-workstations.
  This is actully the advantage of our method, and one of the important information
  delievered by the paper.}

Moreover, regarding the comment below:

{\it On the other hand, the paper lacks important pieces of
  information and many incomplete statements and definitions (see
  below). There is also a theoretical weakness in the sense that the
  main equation used to evaluate the chem. potential seems to be
  derived in an ad hoc fashion, and ``proven numerically'', as the
  authors wrote. It is not easy to rely on an equation/method which
  lacks a solid theoretical background (mostly in statistical
  physics)}

Some of the information asked from the referee is indeed missing,
i.e.~description of the coarse-grained model. \recheck{Actually  this is} a
further advantage of our approach: One does NOT need to derive the
\recheck{coarse-grained models}, but simply takes spherical models interacting via a WCA
potential. It is in fact the thermodynamic force which makes them
compatible with the corresponding full atomistic models. We will add this part in
the Appendix.

However, we tend to strongly disagree on the comment about the lack of
solid theoretical background. \recheck{One can refer to all details of
  the theoretical developement that has alreadly been published in
  Phys.Rev.X.,} and we would like very much to avoid to report again
the whole derivation \recheck{ in this paper, since it would be redundant.}

If, instead, the referee means that he found our paper on Phys.Rev.X
was not solid enough or wrong, then clear scientific arguments must be
brought forward and eventually a comment to the PRX paper should be
done. We did not find helpful the \recheck{words like ``{\bf seems to
    be derived}''}. In the PRX paper, we claimed that the equations
were \recheck{derived \textbf{rigorously} and {\bf exactly}},
\recheck{thus either the referee had found some mistakes in our
  previous paper (then this should be made clear), or he did not read
  the PRX paper.}

In conclusion, we will take several suggestions from {\bf Reviewer 1},
\recheck{however, we will reduce the additions about the
derivation of the formulas to the minimum. Mostly we refer to the exact
equations in the PRX paper}. \recheck{Indeed, we intend} that people focus on
the numerical calculations rather then on the analytical derivations
which can be found in full details in the PRX paper.

We are confident that regarding these points we have your understanding.

\vskip 2cm

Below we report the reply to each referee. The comment of the referees
are in italics.  In order to facilitate the work of the referee, the
additions/modifications in the text will be in red.

@@

{\it ``..... The idea of calculating excess
  chemical potentials from adaptive resolution techniques is appealing}

@@

{\it ``This work is in fact a sort of extension of Ref. 5, and its contribution is just to offer a check of the
statements, hypothesis and calculations made on that previous paper. It is not easy to judge the relevance of this
work because it might contain enough simulations, but not new information. In fact, its main contribution is a table, with comparison with reported data.}

In Ref.5, we claim to have developed a rigorous approach from the
point of view of physics, and gone even further by exploring how far
one could make it mathematically rigorous/exact. We have found that
under the hypothesis:

(1) Thermodynamic limit for the \redc{atomistic and the coarse-grained
  region. (i.e. the number of molecules in both of the two regions are
  infinitely large, and the coarse-grained region is infinitely larger
  than the atomistic region)}

@@

The method is mathematically rigorous and for some aspects even
exact. Conditions (1) and (2) are ideal, so it was important to know
how they can be met in \recheck{practical simulations, especially in
the mixture systems that the solute concentration is dilute.} The aim of this
paper is: {\bf to explore the numerical accuracy of the theory (for
  the chemical potential) proposed in Ref.5}.

We have done it for thirteen representative \recheck{I would remove ``and computationally expensive'', reason as before} systems.
We hope the referee can recognize that this effort is a non trivial one.

{\it On the other hand, the paper lacks important pieces of information and many incomplete statements and definitions (see below). There is also a theoretical
weakness in the sense that the main equation used to evaluate the chem. potential seems to be derived in an ad hoc fashion, and ``proven numerically'', as the authors wrote. It is not easy to rely on an equation/method which lacks a solid theoretical background (mostly in statistical physics)}

The sentence {\bf ``proven numerically''} has been used by the referee
as argument for an {\it ad hoc} formulation of the method, and our
impression is that he/she means that we constructed it based on
\recheck{what we see numerically}. \recheck{ In fact, it is not. In
  Ref.5, the theory was ({\bf explicitly}) derived from thermodynamic
  priciples, and was proved in a rigorous and exact way (in the due
  limits). The conclusion itself does not need further numerical verfication.
  The referee is legitimate to think that it is not rigorous, but
  then he/she should point out explicitly the weak points of Ref.5 and
  provide solid arguments.  However, in this case the referee should 
  correct/comment the paper of Ref.5. The numerical tests
  presented by this manuscript provide extra information on how our
  theory works in practial pure/mixutre systems, in which
  thermodynamic limit can only be approached, but never be reached. We believe this is a plus
  rather than a minus of our work. We are confident that the referee
  sees our point.
}

The sentence {\bf ``proven numerically''} has been used by the referee
as argument for an {\it ad hoc} formulation of the method, and our
impression is that he/she means that we constructed it in order to get
what we wanted. In fact it is not, the formula we have proven
numerically is actually an extra information. We do not need to use
it, in fact we have a technically rigorous way to check the heat given
by the thermostat; this formula just gives an interesting insight,
this is explicitly written in Ref.5. In this paper we have explored it
further and proved that works for other twelve systems. We believe
this is a plus rather than a minus of our work.

In general, as we have said above, we have presented our approach in
Ref.5 ({\bf explicitly}) as rigorous and exact (in the due limits).
The referee is legitimate to think that it is not, but then he/she
should point out explicitly the weak points of Ref.5 and provide solid
arguments.  However, in this case should be corrected/commented the
paper of Ref.5, the numerical result of this paper suggest that at the
simulation level the method cannot be that bad. We are confident that
the referee sees our point.

We found several comments extremely useful to improve the quality of the paper and performed several additional calculations in order to fulfill the request of the referee, we report below the description of all of them.

{\it (1) A description of the cg model is lacking. Please specify how
  do you treat and the interaction potential for cg molecules of
  solvent and solute. I suppose that both types of molecules (solvent
  and solute) free to cross the hybrid layer.  Please confirm. Please
  add a section about the cg model.}

We added it in the appendix. It must be noticed that we actually do
not need to derive a cg model. In fact our approach allows us to use
simple spherical models \recheck{(e.g.~interacting via the WCA potential)} whose size is about the vdW radius of the
specific molecule. It is
then the thermodynamic force that makes the thermodynamics of the cg
models compatible with the atomistic one. Molecules change their
resolution when they pass from one region to another.

{\it (2) Definitions of $V_{AT}$ and $V_{CG}$ are also lacking in terms of particle indexes. The model definition is important to distinguish between total energies and pair-wise terms, e.g. $V_{AT}(xij)$.}

\recheck{we can refer to the new appendix}

{\it (3) Figure 1. ``allows to write an exact Hamiltonian for the atomistic region and thus treat the system in a GC fashion.'' This phrase is
difficult to understand. The AT region is connected with the CG domain (via a transition layer which is not small, but much thicker than the
molecular size). Thus either one has a complete Hamiltonian for the whole AT+HYB+CG system, or else, there is no Hamiltonian. In other words, if
the interactions between AT and HYB, and CG molecules are not derived from H, then there is no Hamiltonian. Also, ``...treat the system in a GC
fashion'' is also difficult to understand (maybe the authors mean ``...consider the AT region as an open subsystem''). I would suggest
deleting this phrase and refer in the text to their theoretical findings in Ref. 5 about particle number distributions p(N),etc.}

@@

In fact we have extended the hybrid region adding a region equal to
the interaction cut off where the resolution is full atomistic, but
such a region is not considered part of the full atomistic region. As
a consequence all the molecules of the atomistic region will interact
with the hybrid region {\bf ONLY} at atomistic level (coarse-grained
molecules are in the region beyond the cut off one). In fact beyond
the cut off radius there is no interaction or better the interaction
is zero. As a consequence the molecules of the atomistic region have a
well defined Hamiltonian. This is the Hamiltonian of an open system,
\recheck{and is parametrized by the the coordinate of the molecules
  that are interacting with the atomistic region}. \recheck{It is comparable with  a
subsystem of a large full atomistic simulation}; the Hamiltonian of the
subsystem is well defined, without any artificial interpolation.  This
is clearly reported in Eq.12 of Ref.5.

Although we believe it is redundant we have reported the explicit form
of the Hamiltonian of the atomistic region also in this paper.

\recheck{{\bf HAN can you write it for me in Latex}}.
\recheck{{\bf See below}}
\begin{align}
  \mathcal H^{\AT}_{N_1} (\vect x_1; \vect x_2, N_2)
  =
  \sum_{i=1}^{N_1}\frac12 m_i\vect v_i^2
  + \sum_{i,j=1}^{N_1}\frac 12 V^\AT_{i,j}
  + \sum_{i=1}^{N_1}\sum_{j=N_1+1}^{N_1+N_2} V^\AT_{i,j},
\end{align}
where $V^\AT_{i,j}$ is the atomistic molecular potential, which is given by
\begin{align}
  V^\AT_{i,j} =
  \sum_{\alpha\in i}\sum_{\beta\in j} V^\AT(\vect r_\alpha - \vect r_\beta).
\end{align}
where index $\alpha$ and $\beta$ denotes the atoms on the
corresponding molecule.  $N_1$ and $N_2$ are the number of molecules
in the AT and atomistic part of the HY region, respectively. $\vect x_1$
and $\vect x_2$ are the phase space variables (joint of velocity and position)
in the  AT and atomistic part of the HY region, respectively.

{\it (4) Page 4. ``...reproduce the first order of the
  probability...'' and ``Higher orders...'' Could the authors specify
  what (small) quantity the order is referred to?, I think it could be
  orders in the density of solute molecules, or maybe in density
  variation with respect the average. Please be more clear when
  referring to ``orders'' elsewhere in the manuscript.}

\recheck{comment from HAN: why not add your short description on the
  orders to the paper? It is concise and would be convenient for the
  readers who are only interested in the numerical result, and do not
  have time to read Ref.5.}

This is clearly described in Ref.5 and we felt it would be redundant
to write it here once again; we have defined the first order of the
probability distribution of a system as its molecular density, a
function in three-dimension. The second order is the two-body
distribution (6-dimensional function) of which the radial distribution
function is a specific expression; the third order is the three-body
distribution (in 9-dimension) etc etc. The full probability is a
3N-dimensional function of which the different orders are a lower
dimensional approximation.

If the referee put this as a condition of rejection, then we would add
it, but we strongly believe that it is not needed for this paper.

{\it (5) Page 5. Auxiliary Hamiltonian. For the sake of clarity, the
  authors should write down this Hamiltonian they refer to, with full
  index details. Also, please write down the form of thermodynamic
  force in the auxiliary Hamiltonian. They mention that simulations
  using the auxiliary Hamiltonian were made with no thermostat. Did
  the authors check that the ``energy'' of this Hamiltonian is
  conserved? Please comment on this important issue.}

\recheck{Comment from HAN: Wo DO use a thermostat for the potential-interpolation
  simulation.  we should be honest on this point. And the prove in the
  appendix also assumes that ths thstem is thermostated. Tell me why you wrote ``
  we do not need to use a thermostat''? Or you simply mean that we in principle we
  do not need, but in practice, to generate the Boltzmann distribution we use it.}

  As before, it is reported in all details in Ref.5 and we thought it
  was redundant. The thermodynamic force must have exactly the same
  numerical formula of the standard AdResS, otherwise it would not
  work for our purposes: if two expressions are formally identical but
  one has the effect of the thermostat and the other not then the
  difference isolates the effect of the thermostat. For the
  Hamiltonian system we use the same formula and impose artificially
  the same density of the AdResS system thus we get a condition on the
  gradient of the molecular density in the transition region.

  Yes, we have checked the conservation of the energy. However this
  was preliminary to everything, in fact it must be clear that if the
  energy was not conserved, then we would also get completely wrong
  results (the heat would either be too low or too high and anyway
  would not converge).

  {\it (6) ``Keeps the density of particles across the system as in
    AdResS''. Do you mean at constant density?}

@@

{\it (7) ``It must be noticed that the relation we obtained form the
  chem. pot. is similar to that obtained in... Ref[6]''. This
  statement is not clear. The equation this method proposes seems not
  similar to that of Ref. 6. In particular the term is quite different
  from which is the corresponding free energy difference appearing in
  Ref. 6. The authors should explicitly write out this essential term
  of the calculation, indicating summations with particle indexes. It
  would seem from the (not-written auxiliary Hamiltonian) that the
  particles index of the first ``w'' is not the same as that of the
  ``nabla w''. Please derive and clarify this term so anybody could
  understand it.

  The effect of discrepancies with Ref.6 are not simple to trace, but
  certainly not zero. The authors state: ``H-Adress and CG-Adress are
  the same up to first order'' (please again, in what?). A reference
  proving this statement is not mentioned, but I guess the authors
  refer to Ref. 7 for that prove. However after reading that paper
  (Ref. 7) I saw no prove of this (just essentially the same phrase,
  without prove). The authors should prove this statement here or
  delete this phrase, or comment about the lack of comparisons, at
  present.}

We see the point of the referee, perhaps we should do a paper with
explicit formulas and data to show similarity and differences between
the two methods.

However since this is not a relevant part of the paper, we decided to
remove any reference to similarities and differences with Ref.6.  We
will cite their second paper dealing with a mixture. We hope that the
referee is happy with the deletion of Ref.6 and addition of a new
reference.

{\it (8) Fig. 2. Please indicate the locations of the HYB domain. A
  difference of about 20 percent in density is not small, although
  *inside* the AT domain, the g(r) might not be too sensible to such
  density difference (as shown in the fig.). On the other hand,
  contrary with what the authors state below, a large transition
  region is beneficial, because molecules have larger space to adapt
  to the change in resolution. Please correct this.}

We disagree with the referee.
\recheck{The large deviation is due to the sampling error, because the
  number of TBA molecules in the system is very small (80, see the Table II).
  We could reach a much better accuracy of the density if we do longer simulations,
  but it would require much more computer time.}
% We could reach a much higher accuracy if
% we set the level of accuracy of the thermodynamic force higher (but
% requires more computer time)
However, we wanted to show that even with 20 percent of (maximum)
difference we get the chemical potential accurate and the density in
the AT region flat.
\recheck{comment from HAN: is the HY region goes
  from 0.5 nm to 3.2 nm, i.e.~the whole range of the figure 2? Then we
  do not show the density in the AT region is flat.}

Regarding the point
that a larger region is beneficial, we must disagree. This is based on
arguments (which we will give next) and long experience with such
systems. What is important in AdResS is that the AT and CG region are
large, in fact if they are large they are going to thermalize and in
general equilibrate substantially the hybrid region.  We have recently
made tests on related issues and found that if one uses a large
thermalized reservoir then small regions where the density or the the
temperature are not targeted, are automatically equilibrated by the
large reservoir. Moreover our analytic derivation suggests the same.
On the contrary a HYB region is charaterized by non physical
description of the potentials and thus it does not have as a reference
any thermodynamic state point. As a consequence the larger the HYB
region the larger the perturbation. Of course there is a minimal size
that is a region equal to the interaction cut off.  These aspects have
been explored and described in details in several previous papers
among which:

(1) Matej Praprotnik, Luigi Delle Site, and Kurt Kremer, .J.Chem.Phys. 123, 224106 (2005).

(2)Silvina Matysiak, Cecilia Clementi, Matej Praprotnik, Kurt Kremer and Luigi Delle Site, J.Chem.Phys. 128, 024503 (2008).

(3) S.Poblete, M.Praprotnik, K.Kremer and L.Delle Site,  J.Chem.Phys. 132, 114101 (2010).

>>>>

\documentclass[12pt]{article}
\usepackage{german}
\usepackage{a4}
\usepackage{graphicx}
\usepackage{pslatex}
\usepackage{color}
\pagestyle{empty}
\parindent0cm
% \addtolength{\textheight}{2cm}
\renewcommand{\sfdefault}{cmss}
\renewcommand{\familydefault}{\sfdefault}
\newcommand{\recheck}[1]{{\color{red} #1}}

\newcommand{\kopf}{\noindent
\raisebox{2.5cm}[0cm][0cm]
{
\parbox[c]{0.88\textwidth}
{
\begin{center}
\textsc{\large Institute for Mathematics, Freie Universit\"{a}t Berlin}\\
\textsc{ PD Dr. Luigi Delle Site}
\end{center}
}
}}
%\raisebox{2cm}[0cm][0cm]{\rule{\textwidth}{0.2mm}}}
\begin{document}
\kopf
\vspace{2cm}
\hfill\raisebox{1.5cm}[0cm][0cm]
{\parbox[t]{0.34\textwidth}
{e-mail:\\
{\small luigi.dellesite@fu-berlin.de}\\
Arnimallee 6\\
D-14195 Berlin\\
Tel:\ +49 (0)30 838 75775 \\
Fax:\ + 49 (0)30 838 75412 \\
}}
\vspace{3cm}
\noindent\raisebox{1.5cm}[0cm][0cm]
{\fbox{\parbox[t]{0.6\textwidth}
{Editor of {\it The Journal of Chemical Physics}\\
}}}
\vspace{1cm}

@@

\vspace{1cm}
\parskip 2ex
Dear Editor

We would like to thank both referees for the criticisms and suggestions.
Below please find a reply to each of the concerns raised by the referees and the
corresponding modification of the manuscript (where required).
In order to facilitate the work of the referees we have highlighted in red, the additions/modifications in the manuscript.

\vskip 1cm
Sincerely,

\newpage

{\bf For the Editor only:}

Regarding the comment of {\bf Reviewer 1}:\\
{\color{blue} \it  It is not easy to judge the relevance of this
  work because it might contain enough simulations, but not new information. In fact, its main contribution is a table, with comparison with reported data.}

In the previous paper published about one year ago in Physical Review X,
we have developed what we have {\bf openly} proposed in that paper as a {\bf rigorous approach} where all the formulas were derived analytically and checked mathematically.
The whole theory is based on two assumptions:

(1) Thermodynamic limit for the coarse-grained region (i.e. number of molecule very large).

@@

In a simulation of course these conditions cannot be met at the ideal level, thus the original idea of this paper was that of treating real systems in order to explore the numerical {\bf accuracy} that the formulas can provide. Thirteen systems (about 300 simulations) were studied with both AdResS and with the thermodynamic integration, this {\it per se} is a major numerical effort, even for large computer facilities, however, the systems were carefully chosen as representative of different classes of liquids and mixtures characterized by different physics. The {\bf new information} consists of thirteen values of the chemical potential which are reproduced by GC-AdResS, thus showing that the method is trustable for such kind of calculations and that the hypothesis (1) and (2) above can be easily met in standard simulations.
In the paper, we have made clear the intention described above, but it seems that the first referee has missed it.
Moreover, the request of the first referee about showing concentration dependent behaviors in order to show that we are consistent with previous work at very low concentrations was actually not of major interest, since we always compare each GC-AdResS calculation with the corresponding TI at the same concentrations. {\bf In our view this was clear also to the referee and it was also clear to him/her the amount of additional computational effort he/she was asking for}.

However, independently from the referee's comment, we decided to do the study about the concentration dependent. In fact, in case of positive outcome, we would show that our method is very robust over different thermodynamic state points (and, as minor point, we would make happy the referee about the consistency with previous work). Indeed our results strongly support such a conclusion. However the computational cost was massive (about 300 additional simulations!), for this reason we ask you to consider this point in case the referee requires additional simulations.

In addition, we did not like the comment of the referee:\\
{\color{blue}\it ``There is also a theoretical
  weakness in the sense that the main equation used to evaluate the chem. potential seems to be derived in an ad hoc fashion, and ``proven numerically'', as the authors wrote.}

It was not the aim of the paper to have an analytic derivation of the formula, this does not mean that the approach is weak and above all the referee gave the impression that we obtained something by manipulating formulas and computational tools in order to ``force'' our results.

Given our major effort over the last 2-3 years to merge analytic/mathematical work with numerical applications to realistic chemical systems, the opinion of the referee is quite discouraging. 
However, our impression is that the first referee did not really read our previous paper and/or he/she may not have the background for catching analytic/mathematical derivations of formulas in full. In any case, in this revised version we have added the analytic derivation of the formula, (and in order to address the point raised by the second referee, we have added also the analytic derivation of the extension of the formula to multicomponent component systems). We have tried to be not too much mathematical, in order to reach the whole community interested in the subject. However, to merge computational results about chemical systems and analytic derivation of the formulas in one coherent paper is not easy, we made a strong effort in such direction and we hope that you will appreciate it.

We have also added two very recent references related to the use of AdResS with the MARTINI force field. The senior authors of such publications pointed out to us the importance of calculating the chemical potential as a test of thermodynamical consistency of the whole system, especially for systems of the kind they are treating (large bio-systems).

\recheck{do not say that...}
Finally, the referee asked us to ``{\it ameliorate} the language; we are fully aware of not being mother tongue Oxford academics, but we asked several colleagues to read the paper and tell us if it was understandable; we got only positive comments, we cannot do more than that.

We have added all the calculations/information asked by the referees and we are confident that now the paper is ready for publication.

\newpage

@@

{\color{blue}{\it ``..... The idea of calculating excess
    chemical potentials from adaptive resolution techniques is appealing}}

@@

{\color{blue}{\it ``This work is in fact a sort of extension of Ref. 5, and its contribution is just to offer a check of the
statements, hypothesis and calculations made on that previous paper. It is not easy to judge the relevance of this
work because it might contain enough simulations, but not new information. In fact, its main contribution is a table, with comparison with reported data.}}

In Ref.5, we claim to have developed a rigorous approach from the
point of view of physics and went even further by exploring how far
one could make it mathematically rigorous/exact. We have found that
under the hypothesis:

(1) Thermodynamic limit for the \recheck{atomistic and the coarse-grained region. (i.e. the number of molecules in both of the two
  regions are infinitely large, and the coarse-grained region is infinitely larger than the atomistic region)}

@@

The method is mathematically rigorous and for some aspects even
exact. Conditions (1) and (2) are ideal, so it was important to know
how they can be met in \recheck{practical simulations, especially in
the mixture systems that the solute concentration is dilute.} The aim of this paper is: {\bf to
  explore the numerical accuracy of the theory (for the chemical
  potential) proposed in Ref.5}.  We have done it for thirteen
representative and computationally expensive systems.  We hope the
referee can recognize that this effort is a non-trivial one. However
as asked by him/her, we have added all the simulations requested (see
below).

{\color{blue}{\it On the other hand, the paper lacks important pieces of information and many incomplete statements and definitions (see below). There is also a theoretical
    weakness in the sense that the main equation used to evaluate the chem. potential seems to be derived in an ad hoc fashion, and ``proven numerically'', as the authors wrote. It is not easy to rely on an equation/method which lacks a solid theoretical background (mostly in statistical physics)}}

% \recheck{
%   The main purpose of the paper was to show the numerical accuracy of the theory, so we did not provide the theoretical derivation of
% }
% We have taken such a comment as a motivation to make an effort and finalize a derivation of the formula defined {\it ``weak''} by the referee. The derivation is presented as one of the results of this paper and has got its own section.
We have taken such a comment as a motivation to \recheck{provide the} derivation of the formula defined {\it ``weak''} by the referee. The derivation is presented as one of the results of this paper and has got its own section.

{\color{blue}{\it (1) A description of the cg model is lacking. Please specify how do you treat and the interaction potential for
cg molecules of solvent and solute. I suppose that both types of molecules (solvent and solute) free to cross the hybrid layer.
Please confirm. Please add a section about the cg model.}}

We added the procedure in the appendix. It must be noticed that, as a
major result of Ref.5, we actually do not need to derive a detailed cg
model. In fact our approach allows us to use simple spherical models
whose size is about the vdW radius of the specific molecule. The spheres
interact via the simple WCA potentials. It is then the thermodynamic force
that makes the thermodynamics of the cg models compatible with the
atomistic one. Both types of molecules change their resolution when they pass from
one region to another.

{\color{blue}{\it (2) Definitions of $V_{AT}$ and $V_{CG}$ are also lacking in terms of particle indexes. The model definition is important to distinguish between total energies and pair-wise terms, e.g. $V_AT(xij)$.}}

Done. See the new Section IV for a detailed description.

{\color{blue}{\it (3) Figure 1. ``allows to write an exact Hamiltonian for the atomistic region and thus treat the system in a GC fashion.'' This phrase is
difficult to understand. The AT region is connected with the CG domain (via a transition layer which is not small, but much thicker than the
molecular size). Thus either one has a complete Hamiltonian for the whole AT+HYB+CG system, or else, there is no Hamiltonian. In other words, if
the interactions between AT and HYB, and CG molecules are not derived from H, then there is no Hamiltonian. Also, ``...treat the system in a GC
fashion'' is also difficult to understand (maybe the authors mean ``...consider the AT region as an open subsystem''). I would suggest
deleting this phrase and refer in the text to their theoretical findings in Ref. 5 about particle number distributions p(N),etc.}}

@@

In fact we have extended the hybrid region adding a region equal to
the interaction cut off where the resolution is full atomistic, but
such a region is not considered part of the full atomistic region. As
a consequence all the molecules of the atomistic region will interact
with the hybrid region {\bf ONLY} at atomistic level (coarse-grained
molecules are in the region beyond the cut off one). In fact beyond
the cut off radius there is no interaction or better the interaction
is zero. As a consequence the molecules of the atomistic region have a
well defined Hamiltonian: The Hamiltonian of an open system.
It is as we take a subsystem of a full atomistic simulation; the Hamiltonian of the subsystem is well defined, without any artificial interpolation.
Moreover an open subsystem of a much larger system is naturally Grand Canonical.
This is clearly reported in Ref.5 and Eq.12 provides the exact definition of the Hamiltonian.
We have added the exact reference for the equation with the specific Hamiltonian, and we do not believe it should be explicitly reported and discussed in the current paper because it would be redundant and anyway it is not crucial to any of the derivations of this work.

{\color{blue}{\it (4) Page 4. ``...reproduce the first order of the
    probability...'' and ``Higher orders...'' Could the authors
    specify what (small) quantity the order is referred to?, I think
    it could be orders in the density of solute molecules, or maybe in
    density variation with respect the average. Please be more clear
    when referring to ``orders'' elsewhere in the manuscript.}}

This is clearly described in Ref.5; we have defined the first order of
the probability distribution of a system as its molecular density, a
function in three-dimensions. The second order is the two-body
distribution (6-dimensional function) of which the radial distribution
function is a specific expression; the third order is the three-body
distribution (in 9 dimensions) etc etc. The full probability is a
3N-dimensional function of which the different orders are a lower
dimensional approximation.

We have added the explicit definition in the revised version of the paper.

{\color{blue}{\it (5) Page 5. Auxiliary Hamiltonian. For the sake of
    clarity, the authors should write down this Hamiltonian they refer
    to, with full index details. Also, please write down the form of
    thermodynamic force in the auxiliary Hamiltonian. They mention
    that simulations using the auxiliary Hamiltonian were made with no
    thermostat. Did the authors check that the ``energy'' of this
    Hamiltonian is conserved? Please comment on this important
    issue.}}

As before, it is reported in all details in Ref.5 and we thought it
was redundant, but we have added it in the revised version. The
thermodynamic force must have exactly the same numerical formula of
the standard AdResS, otherwise it would not work for our purposes: if
two expressions are formally identical but one has the effect of the
thermostat and the other not then the difference isolates the effect
of the thermostat. For the Hamiltonian system we use the same formula
and impose artificially the same density of the AdResS system thus we
get a condition on the gradient of the molecular density in the
transition region, this is now reported in the paper.

Yes, we have checked the conservation of the energy and added a
section dedicated to this issue in the Appendix. There we report the
advantage and limitations of such an approach (i.e. when we can
consider pure conservation and when the coupling to a thermostat is
anyway needed).

{\color{blue}{\it (6) ``Keeps the density of particles across the system as in AdResS''. Do you mean at constant density?}}

@@

{\color{blue}{\it (7) ``It must be noticed that the relation we
    obtained form the chem. pot. is similar to that obtained
    in... Ref[6]''. This statement is not clear. The equation this
    method proposes seems not similar to that of Ref. 6. In particular
    the term is quite different from which is the corresponding free
    energy difference appearing in Ref. 6. The authors should
    explicitly write out this essential term of the calculation,
    indicating summations with particle indexes. It would seem from
    the (not-written auxiliary Hamiltonian) that the particles index
    of the first ``w'' is not the same as that of the ``nabla
    w''. Please derive and clarify this term so anybody could
    understand it.

    The effect of discrepancies with Ref.6 are not
    simple to trace, but certainly not zero. The authors state:
    ``H-Adress and CG-Adress are the same up to first order'' (please
    again, in what?). A reference proving this statement is not
    mentioned, but I guess the authors refer to Ref. 7 for that
    prove. However after reading that paper (Ref. 7) I saw no prove of
    this (just essentially the same phrase, without prove). The
    authors should prove this statement here or delete this phrase, or
    comment about the lack of comparisons, at present.}}

We see the point of the referee, perhaps we should do a paper with explicit formulas and data to show similarity and differences between the two methods.
However since this is not a relevant part of the paper, we decided to remove any reference to similarities and differences with Ref.6.
and add the reference from the same authors on mixtures.

{\color{blue}{\it (8) Fig. 2. Please indicate the locations of the HYB
    domain. A difference of about 20 percent in density is not small,
    although *inside* the AT domain, the g(r) might not be too
    sensible to such density difference (as shown in the fig.). On the
    other hand, contrary with what the authors state below, a large
    transition region is beneficial, because molecules have larger
    space to adapt to the change in resolution. Please correct this.}}

\recheck{The Fig.2 (Fig.4 in the revision) plots the density profile only for the HYB region, as the caption states.}
% We disagree.
\recheck{The relatively large deviation is due to the sampling error, because the
  number of TBA molecules in the system is very small (80, see the Table II, i.e.~the Table IV in the revision).
  We could reach a much better accuracy of the density if we do longer simulations,
  but it would require much more computer time.}
% We could reach a much higher accuracy if we set the level
% of accuracy of the thermodynamic force higher (but requires more
% computer time).
However, we wanted to show that even with 20 percent of
(maximum) difference we get the chemical potential accurate and the
density in the AT region flat.

Regarding the point that a larger
region is beneficial, we must disagree. This is based on arguments
(which we will give next) and long experience with such systems. What
is important in AdResS is that the AT and CG region are large, in fact
if they are large they are going to thermalize and in general
equilibrate substantially the hybrid region.  We have recently made
tests on related issues and found that if one uses a large thermalized
reservoir then small regions where the density or the the temperature
are not targeted, are automatically equilibrated by the large
reservoir. Moreover our analytic derivation suggests the same.  On the
contrary a HYB region is characterized by non-physical description of
the potentials and thus it does not have as a reference any
thermodynamic state point. As a consequence the larger the HYB region
the larger the perturbation. Of course there is a minimal size that is
a region equal to the interaction cut off.  These aspects have been
explored and described in details in several previous papers among
which:
\begin{itemize}
\item (1) Matej Praprotnik, Luigi Delle Site, and Kurt Kremer, .J.Chem.Phys. 123, 224106 (2005).
\item (2)Silvina Matysiak, Cecilia Clementi, Matej Praprotnik, Kurt Kremer and Luigi Delle Site, J.Chem.Phys. 128, 024503 (2008).
\item (3) S.Poblete, M.Praprotnik, K.Kremer and L.Delle Site,  J.Chem.Phys. 132, 114101 (2010).
\end{itemize}
